---
title: "Departmental Survey Summary"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document:
    keep_md: false
---

```{r setup, include=FALSE}
# Global chunk options: hide code, warnings, and messages
knitr::opts_chunk$set(
  echo = FALSE,          # hides code
  warning = FALSE,       # hides warnings
  message = FALSE,       # hides messages
  fig.width = 12,        # default plot width
  fig.height = 8         # default plot height
)

```

## 1. Read and Clean data
```{r echo=FALSE}
library(dplyr)    
library(ggplot2)

d <- read.csv("/Users/oidata/Documents/To macbook - Documents/Kilifi Rural Women Financial Inclusion(30th_Aug_2024)/DPIA/Data governance compliance Survey Responses.xlsx - Data collection questionnaire f.csv")

# Fix column names
names(d) <- gsub("\\.", " ", names(d)) # drops the full stops in column names
names(d) <- sub("\\s+$", "", names(d)) # removes whitespaces

library(carobiner)

# Applies the function to all columns. Makes all column names have title case
d[] <- lapply(d, function(col) fix_name(col, case = "title"))

# d$County.Department <- tools::toTitleCase(tolower(d$County.Department))
d$`County Department` <- ifelse(grepl("land",ignore.case = T,d$`County Department`),"Lands, Energy, Housing, Physical Planning & Urban Development",
                              ifelse(grepl("financ",ignore.case = T,d$`County Department`),"Finance and Economic Planning",
                                     ifelse(grepl("disast",ignore.case = T,d$`County Department`),"Special Programs & Disaster Management (Gender)",d$`County Department`)))

# ----------------------------
# STEP 2: Separate Question Types
# ----------------------------
d <- d[,5:ncol(d)]

# Close-ended variables (categorical/ordinal responses)

close_ended_vars <- c(
  "To what extent are you aware of data protection principles set out under the Data Protection Act  2019",
  "Has everyone in your department  been trained on data protection",
  "Does your County have a data protection policy",
  "Is the data protection policy part of employee human resource obligations",
  "Does your County have a privacy notice",
  "Are members of your department well versed with the county s data protection policy and privacy notice   Only answer if you answered yes to 4 6 above",
  "Does your County have a budget for a data governance programme",
  "If your County has a budget for a data governance programme  how much is it in the current financial year",
  "Is there a designated Data Protection Officer at your County",
  "Are there different levels of access to personal data at your County",
  "Is there a personal data complaints mechanism",
  "How is data hosted at your County",
  "Is data hosted by or on behalf of your County Secure",
  "Do you carry out audits of third parties with whom you share data",
  "Do you have a data retention policy",
  "Is there a data breach response strategy in place",
  "Are there data protection clauses in your contracts",
  "Are there data protection clauses in your County s operational documents",
  "On a scale of 0 to 5  how do you rate your County s data governance compliance   zero being lowest and 5 highest level of compliance"
)

# Closed ended question dataframe
closed <- d[, (names(d) %in% close_ended_vars)]

# Keep all columns EXCEPT the ones in close_ended_vars
open <- d[, !(names(d) %in% close_ended_vars)]
open_ended_vars <- names(open)

```


## 2. What personal data do you process in your department

```{r echo=FALSE}

install.packages(c("tidytext", "dplyr", "tm", "topicmodels", "ggplot2", "textclean", "tidyr"))
library(tidytext)
library(dplyr)
library(tm)
library(topicmodels)
library(ggplot2)
library(textclean)
library(tidyr)

# Read the open variable data
d <- open

# Combine relevant free-text columns into one
# Example: combining responses from one column into a single string
combined_text <- paste(d$`What personal data do you process in your department`, collapse = " ")

# Drop N/A in text analysis
combined_text <- gsub("N/A","",combined_text)

# ----------------------------
# Clean Text with TM
# ----------------------------

text_corpus <- VCorpus(VectorSource(combined_text)) # turns the text into a format R can clean easily (called a corpus)

text_corpus <- text_corpus %>%
  tm_map(content_transformer(tolower)) %>%  # make everything lowercase
  tm_map(removePunctuation) %>%             # remove punctuation
  tm_map(removeNumbers) %>%                 # remove numbers
  tm_map(removeWords, stopwords("en")) %>%  # remove common words like "the", "and", "of"
  tm_map(stripWhitespace)                   # remove extra spaces

# STEP 3: Convert to document-term matrix
dtm <- DocumentTermMatrix(text_corpus)
dtm <- removeSparseTerms(dtm, 0.98)  # remove rare terms

# STEP 4: Run LDA topic modeling
k <- 2  # number of themes to extract
lda_model <- LDA(dtm, k = k, control = list(seed = 1234))

# STEP 5: Get top terms per topic
topics <- tidy(lda_model, matrix = "beta")
top_terms <- topics %>%
  group_by(topic) %>%
  slice_max(beta, n = 10) %>%
  ungroup() %>%
  arrange(topic, -beta)

# STEP 6: Visualize top terms in each theme
top_terms %>%
  mutate(term = reorder_within(term, beta, topic)) %>%
  ggplot(aes(term, beta, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") +
  coord_flip() +
  scale_x_reordered() +
  labs(title = "Top Terms in Detected Themes (Topics)",
       y = "Word Probability", x = NULL)

# Create a simple summary table using data.frame
topic_labels <- data.frame(
  Topic = c("Topic 1", "Topic 2"),
  Label = c("Administrative and Financial Data Handling",
            "Employee Information and Personal Records"),
  Core_Theme = c("Movement and storage of HR, finance, and demographic records",
                 "Data on staff identities, personal info, and internal HR files")
)

print(topic_labels)

```

## 3. What personal sensitive data do you process in your department

```{r echo=FALSE}

# Combining responses from one column into a single string
combined_text <- paste(d$`What personal sensitive data do you process in your department`, collapse = " ")

# Drop N/A in text analysis
combined_text <- gsub("N/A","",combined_text)

# ----------------------------
# Clean Text with TM
# ----------------------------

text_corpus <- VCorpus(VectorSource(combined_text)) # turns the text into a format R can clean easily (called a corpus)

text_corpus <- text_corpus %>%
  tm_map(content_transformer(tolower)) %>%  # make everything lowercase
  tm_map(removePunctuation) %>%             # remove punctuation
  tm_map(removeNumbers) %>%                 # remove numbers
  tm_map(removeWords, stopwords("en")) %>%  # remove common words like "the", "and", "of"
  tm_map(stripWhitespace)                   # remove extra spaces

# STEP 3: Convert to document-term matrix
dtm <- DocumentTermMatrix(text_corpus)
dtm <- removeSparseTerms(dtm, 0.98)  # remove rare terms

# STEP 4: Run LDA topic modeling
k <- 2  # number of themes to extract
lda_model <- LDA(dtm, k = k, method = "Gibbs", control = list(seed = 1234, burnin = 1000, iter = 3000))

# STEP 5: Get top terms per topic
topics <- tidy(lda_model, matrix = "beta")
top_terms <- topics %>%
  group_by(topic) %>%
  slice_max(beta, n = 10) %>%
  ungroup() %>%
  arrange(topic, -beta)

# STEP 6: Visualize top terms in each theme
top_terms %>%
  mutate(term = reorder_within(term, beta, topic)) %>%
  ggplot(aes(term, beta, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") +
  coord_flip() +
  scale_x_reordered() +
  labs(title = "Top Terms in Detected Themes (Topics)",
       y = "Word Probability", x = NULL)

# Create a simple summary table using data.frame
topic_labels <- data.frame(
  Topic = c("Topic 1", "Topic 2"),
  Label = c("Personal Identification, Health & Financial Records",
            "Employee, HR & Administrative Documentation"))
  
print(topic_labels)
# 
# library(knitr)
# 
# kable(topic_labels, caption = "🏷️ Final Topic Labels Summary")
# 
# install.packages("gt")
# library(gt)
# 
# topic_labels %>%
#   gt() %>%
#   tab_header(title = "🏷️ Final Topic Labels Summary")

```

## [4] "What laws provide a legal basis for your County to collect  process  store  and transmit personal data"                                                                                                                 

```{r echo=FALSE}

library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)

# 1. Load responses
responses <- str_to_title(open$`What laws provide a legal basis for your County to collect  process  store  and transmit personal data`)

# 2. # 2. Standardization function

standardize_law <- function(x) {
  x <- gsub("\\.", "", x)  # Escape dot, which is a regex wildcard
  x <- gsub("The Kenyan Constitution 2010|Constitution 2010|Constitution Of Kenya", "Constitution of Kenya 2010", x, ignore.case = TRUE)
  x <- gsub("Dpa 2019|Data Protection Act 2019,", "Data Protection Act 2019", x, ignore.case = TRUE)
  x <- gsub("Office Of The County Attorney Act, 2020", "County Attorney Act 2020", x, ignore.case = TRUE)
  x <- gsub("County Government Act Section 105" , "County Government Act", x, ignore.case = TRUE)
  x <- gsub("Ppda 2015", "Public Procurement & Disposal Act (PPDA) 2015", x, ignore.case = TRUE)
  str_trim(x)
}

# 3. Apply function on responses vector
responses_cleaned <- sapply(responses, standardize_law, USE.NAMES = FALSE)

# 4. Wrap into a dataframe
law_df <- data.frame(response = responses_cleaned, stringsAsFactors = FALSE)

# 5. Remove NAs and unclear responses
law_df <- law_df %>%
  filter(!is.na(response)) %>%
  filter(!str_detect(response, "N/A|Not Sure|No Law At The Moment"))

# 6. Split responses into separate laws
law_split <- law_df %>%
  separate_rows(response, sep = ",| and | And ") %>%   # split by comma or 'and'
  mutate(response = str_trim(response)) %>%            # remove extra whitespace
  filter(response != "")                               # drop empty rows

# 7. Count occurrences of each law
law_counts <- law_split %>%
  count(response, sort = TRUE)

# 8. View final result
law_counts

# 9. Plot as bar chart
ggplot(law_counts, aes(x = reorder(response, n), y = n)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  labs(
    title = "Most Frequently Mentioned Legal Bases for Handling Personal Data",
    x = "Law",
    y = "Mentions"
  ) +
  theme_minimal()

# Visuals per department

# Step 1: Extract relevant columns and filter valid entries

dd <- open %>%
  select(department = `County Department`, law_raw = `What laws provide a legal basis for your County to collect  process  store  and transmit personal data`) %>%
  filter(!is.na(law_raw), !str_detect(tolower(law_raw), "n/a|not sure|no law"))

dd$law_raw <- str_to_title(dd$law_raw)

standardize_law <- function(x) {
  x <- gsub("\\.", "", x)  # Escape dot, which is a regex wildcard
  x <- gsub("The Kenyan Constitution 2010|Constitution 2010|Constitution Of Kenya", "Constitution of Kenya 2010", x, ignore.case = TRUE)
  x <- gsub("Dpa 2019|Data Protection Act 2019,", "Data Protection Act 2019", x, ignore.case = TRUE)
  x <- gsub("Office Of The County Attorney Act, 2020", "County Attorney Act 2020", x, ignore.case = TRUE)
  x <- gsub("County Government Act Section 105" , "County Government Act", x, ignore.case = TRUE)
  x <- gsub("Ppda 2015", "Public Procurement & Disposal Act (PPDA) 2015", x, ignore.case = TRUE)
  str_trim(x)
}

dd$law_raw <- sapply(dd$law_raw, standardize_law, USE.NAMES = FALSE)

# Step 2: Combine all responses per department into a single string

dept_combined <- dd %>%
  group_by(department) %>%
  summarise(all_laws = paste(law_raw, collapse = ", "))

# 3. Split multiple laws into separate rows
dd_long <- dept_combined %>%
  separate_rows(all_laws, sep = ",| and | And ") %>%
  mutate(all_laws = str_trim(all_laws)) %>%
  filter(all_laws != "")

law_counts_by_dept <- dd_long %>%
  count(department, all_laws, sort = TRUE)

law_counts_by_dept

library(ggplot2)

# Plot: Stacked Bar Chart of Laws per Department
ggplot(law_counts_by_dept, aes(x = department, y = n, fill = all_laws)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Legal Bases for Personal Data Processing by Department",
    x = "County Department",
    y = "Mentions of Legal Basis",
    fill = "Law"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(face = "bold", size = 14),
    legend.position = "bottom"
  )


```



## [5] "Who makes decisions on data collection  processing  storage  and transmission"


```{r}

# Load libraries
library(dplyr)
library(tidyr)
library(stringr)
library(tidytext)
library(textstem)
library(ggplot2)

# 1. Load responses
responses <- str_to_title(open$`Who makes decisions on data collection  processing  storage  and transmission`)

# 2. # 2. Standardization function
standardize_law <- function(x) {
  x <- gsub("\\/", " or ", x)  # Escape front slash, which is a regex wildcard
  x <- gsub("Any Person Who Intends To Collect Data", "Enumerators", x, ignore.case = TRUE)
  x <- gsub("Hr","Human Resource Officer", x, ignore.case = TRUE)
  x <- gsub("Chief Officers|Chief Officer - Economic Planning", "Chief Officer", x, ignore.case = TRUE)
  x <- gsub("Accounting Officers" , "Accounting Officer", x, ignore.case = TRUE)
  str_trim(x)
}

# 3. Apply function on responses vector
responses_cleaned <- sapply(responses, standardize_law, USE.NAMES = FALSE)

# 4. Wrap into a dataframe
law_df <- data.frame(response = responses_cleaned, stringsAsFactors = FALSE)

# 5. Remove NAs and unclear responses
law_df <- law_df %>%
  filter(!is.na(response)) %>%
  filter(!str_detect(response, "None"))

# 6. Split responses into separate laws
law_split <- law_df %>%
  separate_rows(response, sep = ",|or ") %>%   # split by comma or 'and'
  mutate(response = str_trim(response)) %>%            # remove extra whitespace
  filter(response != "")                               # drop empty rows

# 7. Count occurrences of each law
law_counts <- law_split %>%
  count(response, sort = TRUE)

# 8. View final result
law_counts

# 9. Plot as bar chart
ggplot(law_counts, aes(x = reorder(response, n), y = n)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  labs(
    title = "Most Frequently Mentioned Decision Makers for Handling Personal Data",
    x = "Decision Makers",
    y = "Mentions"
  ) +
  theme_minimal()

#### Visuals per department

# Step 1: Extract relevant columns and filter valid entries

dd <- open %>%
  select(department = `County Department`, decisionmaker = `Who makes decisions on data collection  processing  storage  and transmission`) %>%
  filter(!is.na(decisionmaker), !str_detect(tolower(decisionmaker), "none"))

dd$decisionmaker <- str_to_title(dd$decisionmaker)

dd$decisionmaker <- sapply(dd$decisionmaker, standardize_law, USE.NAMES = FALSE)

# Step 2: Combine all responses per department into a single string

dept_combined <- dd %>%
  group_by(department) %>%
  summarise(decisionmakers = paste(decisionmaker, collapse = ", "))

# 3. Split multiple laws into separate rows
dd_long <- dept_combined %>%
  separate_rows(decisionmakers, sep = ",|or ") %>%
  mutate(decisionmakers = str_trim(decisionmakers)) %>%
  filter(decisionmakers != "")

law_counts_by_dept <- dd_long %>%
  count(department, decisionmakers, sort = TRUE)

law_counts_by_dept

# Plot: Stacked Bar Chart of Laws per Department
ggplot(law_counts_by_dept, aes(x = department, y = n, fill = decisionmakers)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Decision makers on data collection ,processing ,storage and transmission",
    x = "County Department",
    y = "Mentions of Decision maker",
    fill = "Decision Makers"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),
    plot.title = element_text(face = "bold", size = 14),
    legend.position = "bottom"
  )
```


## [6] "How do you inform data subjects  either county residents  employees  contractors or other  on your collection  processing  storage  or transmission of their data"  

```{r echo=FALSE}

# Combining responses from one column into a single string
combined_text <- paste(d$`How do you inform data subjects  either county residents  employees  contractors or other  on your collection  processing  storage  or transmission of their data`, collapse = " ")

# Drop N/A in text analysis
combined_text <- gsub("N/A","",combined_text)

# ----------------------------
# Clean Text with TM
# ----------------------------

text_corpus <- VCorpus(VectorSource(combined_text)) # turns the text into a format R can clean easily (called a corpus)

text_corpus <- text_corpus %>%
  tm_map(content_transformer(tolower)) %>%  # make everything lowercase
  tm_map(removePunctuation) %>%             # remove punctuation
  tm_map(removeNumbers) %>%                 # remove numbers
  tm_map(removeWords, stopwords("en")) %>%  # remove common words like "the", "and", "of"
  tm_map(stripWhitespace)                   # remove extra spaces

# STEP 3: Convert to document-term matrix
dtm <- DocumentTermMatrix(text_corpus)
dtm <- removeSparseTerms(dtm, 0.98)  # remove rare terms

# STEP 4: Run LDA topic modeling
k <- 2  # number of themes to extract
lda_model <- LDA(dtm, k = k, control = list(seed = 1234))

# STEP 5: Get top terms per topic
topics <- tidy(lda_model, matrix = "beta")
top_terms <- topics %>%
  group_by(topic) %>%
  slice_max(beta, n = 10) %>%
  ungroup() %>%
  arrange(topic, -beta)

# STEP 6: Visualize top terms in each theme
top_terms %>%
  mutate(term = reorder_within(term, beta, topic)) %>%
  ggplot(aes(term, beta, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") +
  coord_flip() +
  scale_x_reordered() +
  labs(title = "Top Terms in Detected Themes (Topics)",
       y = "Word Probability", x = NULL)

# Load necessary packages
library(tibble)
library(gt)  # install.packages("gt") if not installed

# Create summary of themes
summary_themes <- tibble::tibble(
  Topic = c("1", "2"),
  Theme = c(
    "Public Participation & Traditional Channels",
    "Formal Internal & Written Communication"
  ),
  `Communication Approach` = c(
    "Barazas, radio, calls, emails, public forums",
    "Letters, departmental memos, confidentiality"
  )
)

# Print the tibble
print(summary_themes)

```

### Summary
#### Public Participation & Traditional Communication Channels


 - This topic suggests that some departments rely on public forums (barazas), community outreach (radio, public participation), and direct communication (phone calls, emails, correspondence) to inform citizens. These methods are more inclusive and accessible, especially in rural or low-digital-access areas.

#### Formal Written & Departmental Communication

 - This topic emphasizes written communication (letters, mails, correspondence), internal awareness (divisional, departmental), and concepts of confidentiality, consent and enlightenment. It suggests a more structured and perhaps internal-focused approach to data subject communication.

 - These themes reflect a dual strategy by reaching citizens via community engagement tools and internal stakeholders (employees, contractors) through formal notices and memos.
 
 
### [7] "What are the primary data collection methods at your County"                                                                                                                                                            

```{r echo=FALSE}

library(stringr)

# 1. Load responses
responses <- str_to_title(open$`What are the primary data collection methods at your County`)

# 2. Standardization function

standardize_law <- function(x) {
  x <- gsub("\\/|\\#|\\?|\\*|\\.", "", x)
  x <- gsub("Questionare|Questionnaire|Questionares|Questionnaires", "Questionnaires", ignore.case = TRUE, x)
  x <- gsub("Discussion|Discussions","Discussions",ignore.case = TRUE,x)
  x <- gsub("focused","Focus",ignore.case = TRUE, x)
  x <- gsub("and",",",ignore.case = TRUE,x)
  x <- gsub("\\s*,\\s*|,\\s*", ",", x)
  str_trim(x)
}

# 3. Apply function on responses vector

clean_responses <- function(responses) {
  responses_cleaned <- sapply(responses, standardize_law, USE.NAMES = FALSE)
  
  # Manual corrections
  responses_cleaned[1]  <- NA
  responses_cleaned[2]  <- "Questionnaires,Observation,Interviews"
  responses_cleaned[3]  <- "Questionnaires,Interviews"
  responses_cleaned[6]  <- "Questionnaires,Interviews"
  responses_cleaned[8]  <- "Questionnaires,Surveys,Consultative Meetings,Focus Group Discussions"
  responses_cleaned[15] <- "Surveys,Surveys"
  
  # Trim whitespace
  responses_cleaned <- str_trim(responses_cleaned)
  
  return(responses_cleaned)
}

# 4. Wrap into a dataframe
law_df <- data.frame(response = responses_cleaned, stringsAsFactors = FALSE)

# 5. Remove NAs and unclear responses
law_df <- law_df %>%
  filter(!is.na(responses_cleaned))

# 6. Split responses into separate laws
law_split <- law_df %>%
  separate_rows(response, sep = ",") %>%        # split by comma
  mutate(response = str_trim(response)) %>%     # remove extra whitespace
  filter(response != "")                        # drop empty rows

# 7. Count occurrences of each primary data collection method
law_counts <- law_split %>%
  count(response, sort = TRUE)

# 8. View final result
law_counts

# 9. Plot as bar chart
ggplot(law_counts, aes(x = reorder(response, n), y = n)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  labs(
    title = "Most Frequently Mentioned Data collection methods",
    x = "Method type",
    y = "Mentions"
  ) +
  theme_minimal()

#### Visuals per department

# Step 1: Extract relevant columns and filter valid entries
dd <- open %>%
  select(
    department = `County Department`,method = `What are the primary data collection methods at your County`) %>%
  filter(!is.na(method)) %>%               # Filter out NA values
  mutate(method = str_to_title(method))    # Clean method column

dd$method <- clean_responses(dd$method)

dd <- dd %>%
  filter(!is.na(method)) %>%
  filter(method != "Not Sure")

dept_combined <- dd %>%
  group_by(department) %>%
  summarise(method = paste(method, collapse = ", "))

# 3. Split multiple laws into separate rows
dd_long <- dept_combined %>%
  separate_rows(method, sep = ",") %>%
  mutate(method = str_trim(method)) %>%
  filter(method != "")

law_counts_by_dept <- dd_long %>%
  count(department, method, sort = TRUE)

law_counts_by_dept

# Plot: Stacked Bar Chart of Laws per Department
ggplot(law_counts_by_dept, aes(x = department, y = n, fill = method)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Primary data collection methods",
    x = "County Department",
    y = "Mentions",
    fill = "Method type"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),
    plot.title = element_text(face = "bold", size = 14),
    legend.position = "bottom"
  )

```

###  [8] "What are the primary data storage methods at your County"                                                                                                                                                               

```{r echo=FALSE}

# Count feedback categories

category_counts <- d %>%
  group_by(`County Department`, `Are members of your department well versed with the county s data protection policy and privacy notice   Only answer if you answered yes to 4 6 above`) %>%
  summarise(total_responses = n()) %>%
  mutate(proportion = total_responses / sum(total_responses))

category_counts

# Stacked bar chart for categorical feedback

ggplot(category_counts, aes(x = `County Department`, y = total_responses, fill = `Are members of your department well versed with the county s data protection policy and privacy notice   Only answer if you answered yes to 4 6 above`)) +
  geom_bar(stat = "identity") +
  labs(title = "Are members of your department well versed with the county s data protection policy and privacy notice   Only answer if you answered yes to 4 6 above") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))


# contingency table to show feedback distribution per department 

table(d$`County Department`,d$`Are members of your department well versed with the county s data protection policy and privacy notice   Only answer if you answered yes to 4 6 above`)

# % distribution of responses
response <- d %>%
  group_by(`Are members of your department well versed with the county s data protection policy and privacy notice   Only answer if you answered yes to 4 6 above`) %>%
  summarize(total_responses = n()) %>%
  mutate(proportion = (total_responses / sum(total_responses))*100)

response

ggplot(response, aes(x = `Are members of your department well versed with the county s data protection policy and privacy notice   Only answer if you answered yes to 4 6 above`, y = proportion, fill = `Are members of your department well versed with the county s data protection policy and privacy notice   Only answer if you answered yes to 4 6 above`)) +
  geom_bar(stat = "identity") +
  labs(title = "Response type proportion (%)") +
  theme_minimal() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# # export the table to excel
# write.csv(f,file = "Feedback per Department of how aware people are of data protection principles")
```

###  [9] "What are the uses of data at your County"                                                                                                                                                                               

```{r echo=FALSE}
# Count feedback categories

category_counts <- d %>%
  group_by(`County Department`, `Does your County have a budget for a data governance programme`) %>%
  summarise(total_responses = n()) %>%
  mutate(proportion = total_responses / sum(total_responses))

category_counts

# Stacked bar chart for categorical feedback

ggplot(category_counts, aes(x = `County Department`, y = total_responses, fill = `Does your County have a budget for a data governance programme`)) +
  geom_bar(stat = "identity") +
  labs(title = "Does your County have a budget for a data governance programme") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# contingency table to show feedback distribution per department 

table(d$`County Department`,d$`Does your County have a budget for a data governance programme`)

# % distribution of responses
response <- d %>%
  group_by(`Does your County have a budget for a data governance programme`) %>%
  summarize(total_responses = n()) %>%
  mutate(proportion = (total_responses / sum(total_responses))*100)

response

ggplot(response, aes(x = `Does your County have a budget for a data governance programme`, y = proportion, fill = `Does your County have a budget for a data governance programme`)) +
  geom_bar(stat = "identity") +
  labs(title = "Response type proportion (%)") +
  theme_minimal() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# # export the table to excel
# write.csv(f,file = "Feedback per Department of how aware people are of data protection principles")
```


### 8. If your County has a budget for a data governance programme  how much is it in the current financial year                                         
                                                            
```{r echo=FALSE}

# drop NA and Nil entries
unique(d$`If your County has a budget for a data governance programme  how much is it in the current financial year`)

d$`If your County has a budget for a data governance programme  how much is it in the current financial year` <- 
  ifelse(d$`If your County has a budget for a data governance programme  how much is it in the current financial year` %in% c("N/a","Nil"), NA,
         ifelse(d$`If your County has a budget for a data governance programme  how much is it in the current financial year` == "I Don't Know", "Not Sure",d$`If your County has a budget for a data governance programme  how much is it in the current financial year`))

unique(d$`If your County has a budget for a data governance programme  how much is it in the current financial year`)

# Count feedback categories

category_counts <- d %>%
  group_by(`County Department`, `If your County has a budget for a data governance programme  how much is it in the current financial year`) %>%
  summarise(total_responses = n()) %>%
  mutate(proportion = total_responses / sum(total_responses))

category_counts

# Stacked bar chart for categorical feedback

ggplot(category_counts, aes(x = `County Department`, y = total_responses, fill = `If your County has a budget for a data governance programme  how much is it in the current financial year`)) +
  geom_bar(stat = "identity") +
  labs(title = "If your County has a budget for a data governance programme  how much is it in the current financial year") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))


# contingency table to show feedback distribution per department 

table(d$`County Department`,d$`If your County has a budget for a data governance programme  how much is it in the current financial year`)

# % distribution of responses
response <- d %>%
  group_by(`If your County has a budget for a data governance programme  how much is it in the current financial year`) %>%
  summarize(total_responses = n()) %>%
  mutate(proportion = (total_responses / sum(total_responses))*100)

response

ggplot(response, aes(x = `If your County has a budget for a data governance programme  how much is it in the current financial year`, y = proportion, fill = `If your County has a budget for a data governance programme  how much is it in the current financial year`)) +
  geom_bar(stat = "identity") +
  labs(title = "Response type proportion (%)") +
  theme_minimal() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# # export the table to excel
# write.csv(f,file = "Feedback per Department of how aware people are of data protection principles")
```

### 9. Is there a designated Data Protection Officer at your County                                        
                                                            
```{r echo=FALSE}
# Count feedback categories

category_counts <- d %>%
  group_by(`County Department`, `Is there a designated Data Protection Officer at your County`) %>%
  summarise(total_responses = n()) %>%
  mutate(proportion = total_responses / sum(total_responses))

category_counts

# Stacked bar chart for categorical feedback

ggplot(category_counts, aes(x = `County Department`, y = total_responses, fill = `Is there a designated Data Protection Officer at your County`)) +
  geom_bar(stat = "identity") +
  labs(title = "Is there a designated Data Protection Officer at your County") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# contingency table to show feedback distribution per department 

table(d$`County Department`,d$`Is there a designated Data Protection Officer at your County`)

# % distribution of responses
response <- d %>%
  group_by(`Is there a designated Data Protection Officer at your County`) %>%
  summarize(total_responses = n()) %>%
  mutate(proportion = (total_responses / sum(total_responses))*100)

response

ggplot(response, aes(x = `Is there a designated Data Protection Officer at your County`, y = proportion, fill = `Is there a designated Data Protection Officer at your County`)) +
  geom_bar(stat = "identity") +
  labs(title = "Response type proportion (%)") +
  theme_minimal() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# # export the table to excel
# write.csv(f,file = "Feedback per Department of how aware people are of data protection principles")
```

### 10. [11] Are there different levels of access to personal data at your County                                                                                

```{r echo=FALSE}
# Count feedback categories

category_counts <- d %>%
  group_by(`County Department`, `Are there different levels of access to personal data at your County`) %>%
  summarise(total_responses = n()) %>%
  mutate(proportion = total_responses / sum(total_responses))

category_counts

# Stacked bar chart for categorical feedback

ggplot(category_counts, aes(x = `County Department`, y = total_responses, fill = `Are there different levels of access to personal data at your County`)) +
  geom_bar(stat = "identity") +
  labs(title = "Are there different levels of access to personal data at your County") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# contingency table to show feedback distribution per department 

table(d$`County Department`,d$`Are there different levels of access to personal data at your County`)

# % distribution of responses
response <- d %>%
  group_by(`Are there different levels of access to personal data at your County`) %>%
  summarize(total_responses = n()) %>%
  mutate(proportion = (total_responses / sum(total_responses))*100)

response

ggplot(response, aes(x = `Are there different levels of access to personal data at your County`, y = proportion, fill = `Are there different levels of access to personal data at your County`)) +
  geom_bar(stat = "identity") +
  labs(title = "Response type proportion (%)") +
  theme_minimal() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# # export the table to excel
# write.csv(f,file = "Feedback per Department of how aware people are of data protection principles")
```


### 11. [12] Is there a personal data complaints mechanism                                                                                                       

```{r echo=FALSE}
# Count feedback categories

category_counts <- d %>%
  group_by(`County Department`, `Is there a personal data complaints mechanism`) %>%
  summarise(total_responses = n()) %>%
  mutate(proportion = total_responses / sum(total_responses))

category_counts

# Stacked bar chart for categorical feedback

ggplot(category_counts, aes(x = `County Department`, y = total_responses, fill = `Is there a personal data complaints mechanism`)) +
  geom_bar(stat = "identity") +
  labs(title = "Is there a personal data complaints mechanism") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# contingency table to show feedback distribution per department 

table(d$`County Department`,d$`Is there a personal data complaints mechanism`)

# % distribution of responses
response <- d %>%
  group_by(`Is there a personal data complaints mechanism`) %>%
  summarize(total_responses = n()) %>%
  mutate(proportion = (total_responses / sum(total_responses))*100)

response

ggplot(response, aes(x = `Is there a personal data complaints mechanism`, y = proportion, fill = `Is there a personal data complaints mechanism`)) +
  geom_bar(stat = "identity") +
  labs(title = "Response type proportion (%)") +
  theme_minimal() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# # export the table to excel
# write.csv(f,file = "Feedback per Department of how aware people are of data protection principles")
```

### 12. [13] "How is data hosted at your County"                                                                                                                    

```{r echo=FALSE}
# Count feedback categories

category_counts <- d %>%
  group_by(`County Department`, `How is data hosted at your County`) %>%
  summarise(total_responses = n()) %>%
  mutate(proportion = total_responses / sum(total_responses))

category_counts

# Stacked bar chart for categorical feedback

ggplot(category_counts, aes(x = `County Department`, y = total_responses, fill = `How is data hosted at your County`)) +
  geom_bar(stat = "identity") +
  labs(title = "How is data hosted at your County") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# contingency table to show feedback distribution per department 

table(d$`County Department`,d$`How is data hosted at your County`)

# % distribution of responses
response <- d %>%
  group_by(`How is data hosted at your County`) %>%
  summarize(total_responses = n()) %>%
  mutate(proportion = (total_responses / sum(total_responses))*100)

response

ggplot(response, aes(x = `How is data hosted at your County`, y = proportion, fill = `How is data hosted at your County`)) +
  geom_bar(stat = "identity") +
  labs(title = "Response type proportion (%)") +
  theme_minimal() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# # export the table to excel
# write.csv(f,file = "Feedback per Department of how aware people are of data protection principles")
```


### 13. [14] "Is data hosted by or on behalf of your County Secure"                                                                                                 

```{r echo=FALSE}

unique(d$`Is data hosted by or on behalf of your County Secure`)

d$`Is data hosted by or on behalf of your County Secure` <- ifelse(d$`Is data hosted by or on behalf of your County Secure` %in% c("Maybe","I Don't Know"),"Not Sure",d$`Is data hosted by or on behalf of your County Secure`)

unique(d$`Is data hosted by or on behalf of your County Secure`)

# Count feedback categories

category_counts <- d %>%
  group_by(`County Department`, `Is data hosted by or on behalf of your County Secure`) %>%
  summarise(total_responses = n()) %>%
  mutate(proportion = total_responses / sum(total_responses))

category_counts

# Stacked bar chart for categorical feedback

ggplot(category_counts, aes(x = `County Department`, y = total_responses, fill = `Is data hosted by or on behalf of your County Secure`)) +
  geom_bar(stat = "identity") +
  labs(title = "Is data hosted by or on behalf of your County Secure") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# contingency table to show feedback distribution per department 

table(d$`County Department`,d$`Is data hosted by or on behalf of your County Secure`)

# % distribution of responses
response <- d %>%
  group_by(`Is data hosted by or on behalf of your County Secure`) %>%
  summarize(total_responses = n()) %>%
  mutate(proportion = (total_responses / sum(total_responses))*100)

response

ggplot(response, aes(x = `Is data hosted by or on behalf of your County Secure`, y = proportion, fill = `Is data hosted by or on behalf of your County Secure`)) +
  geom_bar(stat = "identity") +
  labs(title = "Response type proportion (%)") +
  theme_minimal() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# # export the table to excel
# write.csv(f,file = "Feedback per Department of how aware people are of data protection principles")
```


### 14. [15] "Do you carry out audits of third parties with whom you share data"                                                                                    
                                                                            
```{r echo=FALSE}

unique(d$`Do you carry out audits of third parties with whom you share data`)

d$`Do you carry out audits of third parties with whom you share data` <- ifelse(d$`Do you carry out audits of third parties with whom you share data` %in% c("I Don't Know","Maybe"), "Not Sure",d$`Do you carry out audits of third parties with whom you share data`)

# Count feedback categories

category_counts <- d %>%
  group_by(`County Department`, `Do you carry out audits of third parties with whom you share data`) %>%
  summarise(total_responses = n()) %>%
  mutate(proportion = total_responses / sum(total_responses))

category_counts

# Stacked bar chart for categorical feedback

ggplot(category_counts, aes(x = `County Department`, y = total_responses, fill = `Do you carry out audits of third parties with whom you share data`)) +
  geom_bar(stat = "identity") +
  labs(title = "Do you carry out audits of third parties with whom you share data") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# contingency table to show feedback distribution per department 

table(d$`County Department`,d$`Do you carry out audits of third parties with whom you share data`)

# % distribution of responses
response <- d %>%
  group_by(`Do you carry out audits of third parties with whom you share data`) %>%
  summarize(total_responses = n()) %>%
  mutate(proportion = (total_responses / sum(total_responses))*100)

response

ggplot(response, aes(x = `Do you carry out audits of third parties with whom you share data`, y = proportion, fill = `Do you carry out audits of third parties with whom you share data`)) +
  geom_bar(stat = "identity") +
  labs(title = "Response type proportion (%)") +
  theme_minimal() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# # export the table to excel
# write.csv(f,file = "Feedback per Department of how aware people are of data protection principles")
```

### 15. [16] "Do you have a data retention policy"                                                                                                                  
                                                                            
```{r echo=FALSE}

# Count feedback categories

category_counts <- d %>%
  group_by(`County Department`, `Do you have a data retention policy`) %>%
  summarise(total_responses = n()) %>%
  mutate(proportion = total_responses / sum(total_responses))

category_counts

# Stacked bar chart for categorical feedback

ggplot(category_counts, aes(x = `County Department`, y = total_responses, fill = `Do you have a data retention policy`)) +
  geom_bar(stat = "identity") +
  labs(title = "Do you have a data retention policy") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# contingency table to show feedback distribution per department 

table(d$`County Department`,d$`Do you have a data retention policy`)

# % distribution of responses
response <- d %>%
  group_by(`Do you have a data retention policy`) %>%
  summarize(total_responses = n()) %>%
  mutate(proportion = (total_responses / sum(total_responses))*100)

response

ggplot(response, aes(x = `Do you have a data retention policy`, y = proportion, fill = `Do you have a data retention policy`)) +
  geom_bar(stat = "identity") +
  labs(title = "Response type proportion (%)") +
  theme_minimal() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# # export the table to excel
# write.csv(f,file = "Feedback per Department of how aware people are of data protection principles")
```

### 16. [17] "Is there a data breach response strategy in place"                                                                                                    
                                                                                                        
```{r echo=FALSE}

# Count feedback categories

category_counts <- d %>%
  group_by(`County Department`, `Is there a data breach response strategy in place`) %>%
  summarise(total_responses = n()) %>%
  mutate(proportion = total_responses / sum(total_responses))

category_counts

# Stacked bar chart for categorical feedback

ggplot(category_counts, aes(x = `County Department`, y = total_responses, fill = `Is there a data breach response strategy in place`)) +
  geom_bar(stat = "identity") +
  labs(title = "Is there a data breach response strategy in place") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# contingency table to show feedback distribution per department 

table(d$`County Department`,d$`Is there a data breach response strategy in place`)

# % distribution of responses
response <- d %>%
  group_by(`Is there a data breach response strategy in place`) %>%
  summarize(total_responses = n()) %>%
  mutate(proportion = (total_responses / sum(total_responses))*100)

response

ggplot(response, aes(x = `Is there a data breach response strategy in place`, y = proportion, fill = `Is there a data breach response strategy in place`)) +
  geom_bar(stat = "identity") +
  labs(title = "Response type proportion (%)") +
  theme_minimal() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# # export the table to excel
# write.csv(f,file = "Feedback per Department of how aware people are of data protection principles")
```

### 17. [18] "Are there data protection clauses in your contracts"                                                                                                  
                                                                                                    
                                                                                                        
```{r echo=FALSE}

# Count feedback categories

category_counts <- d %>%
  group_by(`County Department`, `Are there data protection clauses in your contracts`) %>%
  summarise(total_responses = n()) %>%
  mutate(proportion = total_responses / sum(total_responses))

category_counts

# Stacked bar chart for categorical feedback

ggplot(category_counts, aes(x = `County Department`, y = total_responses, fill = `Are there data protection clauses in your contracts`)) +
  geom_bar(stat = "identity") +
  labs(title = "Are there data protection clauses in your contracts") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# contingency table to show feedback distribution per department 

table(d$`County Department`,d$`Are there data protection clauses in your contracts`)

# % distribution of responses
response <- d %>%
  group_by(`Are there data protection clauses in your contracts`) %>%
  summarize(total_responses = n()) %>%
  mutate(proportion = (total_responses / sum(total_responses))*100)

response

ggplot(response, aes(x = `Are there data protection clauses in your contracts`, y = proportion, fill = `Are there data protection clauses in your contracts`)) +
  geom_bar(stat = "identity") +
  labs(title = "Response type proportion (%)") +
  theme_minimal() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# # export the table to excel
# write.csv(f,file = "Feedback per Department of how aware people are of data protection principles")
```

### 18. [19] "Are there data protection clauses in your County s operational documents"
                                                                                                    
                                                                                                        
```{r echo=FALSE}

# Count feedback categories

category_counts <- d %>%
  group_by(`County Department`, `Are there data protection clauses in your County s operational documents`) %>%
  summarise(total_responses = n()) %>%
  mutate(proportion = total_responses / sum(total_responses))

category_counts

# Stacked bar chart for categorical feedback

ggplot(category_counts, aes(x = `County Department`, y = total_responses, fill = `Are there data protection clauses in your County s operational documents`)) +
  geom_bar(stat = "identity") +
  labs(title = "Are there data protection clauses in your County s operational documents") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# contingency table to show feedback distribution per department 

table(d$`County Department`,d$`Are there data protection clauses in your County s operational documents`)

# % distribution of responses
response <- d %>%
  group_by(`Are there data protection clauses in your County s operational documents`) %>%
  summarize(total_responses = n()) %>%
  mutate(proportion = (total_responses / sum(total_responses))*100)

response

ggplot(response, aes(x = `Are there data protection clauses in your County s operational documents`, y = proportion, fill = `Are there data protection clauses in your County s operational documents`)) +
  geom_bar(stat = "identity") +
  labs(title = "Response type proportion (%)") +
  theme_minimal() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# # export the table to excel
# write.csv(f,file = "Feedback per Department of how aware people are of data protection principles")
```

### 19. [20] "On a scale of 0 to 5  how do you rate your County s data governance compliance   zero being lowest and 5 highest level of compliance"                 
                                                                                                    
```{r echo=FALSE}

# Count feedback categories

category_counts <- d %>%
  group_by(`County Department`, `On a scale of 0 to 5  how do you rate your County s data governance compliance   zero being lowest and 5 highest level of compliance`) %>%
  summarise(total_responses = n()) %>%
  mutate(proportion = total_responses / sum(total_responses))

category_counts

# Stacked bar chart for categorical feedback

ggplot(category_counts, aes(x = `County Department`, y = total_responses, fill = `On a scale of 0 to 5  how do you rate your County s data governance compliance   zero being lowest and 5 highest level of compliance`)) +
  geom_bar(stat = "identity") +
  labs(title = "On a scale of 0 to 5  how do you rate your County s data governance compliance   zero being lowest and 5 highest level of compliance") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + 
  coord_flip()

# contingency table to show feedback distribution per department 

table(d$`County Department`,d$`On a scale of 0 to 5  how do you rate your County s data governance compliance   zero being lowest and 5 highest level of compliance`)

# % distribution of responses
response <- d %>%
  group_by(`On a scale of 0 to 5  how do you rate your County s data governance compliance   zero being lowest and 5 highest level of compliance`) %>%
  summarize(total_responses = n()) %>%
  mutate(proportion = (total_responses / sum(total_responses))*100)

response

ggplot(response, aes(x = `On a scale of 0 to 5  how do you rate your County s data governance compliance   zero being lowest and 5 highest level of compliance`, y = proportion, fill = `On a scale of 0 to 5  how do you rate your County s data governance compliance   zero being lowest and 5 highest level of compliance`)) +
  geom_bar(stat = "identity") +
  labs(title = "Response type proportion (%)") +
  theme_minimal() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# # export the table to excel
# write.csv(f,file = "Feedback per Department of how aware people are of data protection principles")
```









